---
title: "ESM 244 Lab 7"
author: "Sam Csik"
date: "2/20/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tmap) 
library(sf) # works beautifully for polygons
library(spatstat)
library(maptools)
library(sp)
library(raster)
library(gstat)
```

Part 1. Hawaii raster practice
```{r}
hi_par <- raster("PAR_CLIM_M.tif")
plot(hi_par)

hi_sst <- raster("SST_LTM.tif")
plot(hi_sst)

hi_chl <- raster("CHL_LTM.tif")
plot(hi_chl)

par(mfrow = c(1,3))
plot(hi_par)
plot(hi_sst)
plot(hi_chl)
```

Reproject to WGS84:
```{r}
wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

# reproject raster data
hi_sst_84 <- projectRaster(hi_sst, crs = wgs84, method = "bilinear")
# look at projection. seems good!
hi_sst_84@crs
hi_sst_84@extent # lat and long data
```

raster::aggregate() for resampling
```{r}
sst_rs <- aggregate(hi_sst_84, fact = 10)
plot(sst_rs) # resolution has decreased by factor of 10; good for testing data since high-level raster info can take a really long time to load
```

crop a raster
```{r}
hi_sst_84@extent
plot(hi_sst_84) # uncropped

bounds <- as(extent(-156.2, -154.5, 18.7, 20.4), 'SpatialPolygons') # as() allows us later on to specify somehting as a spatial polygon

crs(bounds) <- crs(hi_sst_84) # match the raster for hi_sst_84?? i missed this part

sst_crop <- crop(hi_sst_84, bounds) # crop hi_sst_84 raster by the polygon bounds
plot(sst_crop)
```

Simple raster math:

Nonsensical variable called "tropicality" that is the sum of PAR + SST + 2*ChlA, and we want to map that variable...

NOTE: for thematic (categorical) raster data, consider using method "ngm"
```{r}
hi_par_84 <- projectRaster(hi_par, crs = wgs84, metho = "bilinear")

hi_chla_84 <- projectRaster(hi_chl, crs = wgs84, method = "bilinear")

par(mfrow = c(1,3))
```

```{r}
trop <- hi_par_84 + hi_sst_84 + hi_chla_84
plot(trop)
```

Now let's try to look at something using tmap:
```{r}
islands <- read_sf(dsn = "islands", layer = "Island_boundaries") %>% 
  dplyr::select(Island) %>% 
  st_simplify(dTolerance = 10) %>% 
  st_transform(crs = 4326)
  
plot(islands)
```

```{r}
tmap_mode("plot")

sst_map <- tm_shape(hi_sst_84) + tm_raster(title = "Mean Sea Surface Temperature") +
  tm_layout(bg.color = "navyblue",
            legend.position = c("left", "bottom"),
            legend.text.color = "white",
            legend.text.size = 0.5) +
  tm_shape(islands) +
  tm_fill("darkgreen") 

tmap_save(sst_map, "sam_hawaii.png", height = 5)
```

Conditional rasters and masking

Let's say: we have a sensitive species and we are trying to find habitat that it might like. We know: they like warm water (average SST >= 25.6 degrees) and solar radiation (PAR) below 54.

```{r}
extent(hi_sst_84) <- extent(hi_par_84) # make extents match across rasters

compareRaster(sst_new, hi_par_84) # rasters don't match yet

cr <- raster(nrow = 822,
             ncol = 1229,
             xmn = -160.4365,
             xmx = -154.5373,
             ymn = 18.7309,
             ymx = 22.44634) # creates raster with same number of rows, columns, and extent as hi_sst_84

sst_new <- resample(hi_sst_84, cr, method = "bilinear")

compareRaster(sst_new, hi_par_84) # now they match!

plot(sst_new)
plot(hi_par_84)

# make a cropped version just for Kauai
bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')

crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(par_kauai)
plot(sst_kauai)
```

Now we only want to isolate regions where temp >=25.4, PAR < 54
```{r}
par_hab <- par_kauai
par_hab[par_hab >= 54] <- NA
plot(par_hab)

sst_hab <- sst_kauai
sst_hab[sst_hab < 25.5] <- NA
plot(sst_hab)

# where do these overlap? raster::mask

suit_hab <- mask(sst_hab, par_hab)
plot(suit_hab)
```

